---
output: html_document
---

In my [previous blog on 30 Sep 2016][2], I had just discovered the plotly package and was blown away by how easy it was to use plotly to make ggplot2 plots reactive to the user's mouse. 

Well I've just stumbled across another package to achieve this called [ggiraph][1], **and I ggiraph like even more!!**

In my opinion [ggiraph][1] integrates with ggolot2 in a nicer way than plotly to achieve basic reactivity to the user's mouse. 

Below is some code to make same fake data as I used in my last blog post, but this time I use [ggiraph][1] to turn the static ggplot2 plot from my last blog post into a plot with user reactivity.

The code below uses functions from the "[ggplot2](https://cran.r-project.org/web/packages/ggplot2/ggplot2.pdf)", "[stringr](https://cran.r-project.org/web/packages/stringr/stringr.pdf)",  "[dplyr](https://cran.r-project.org/web/packages/dplyr/dplyr.pdf)", "[knitr](https://cran.r-project.org/web/packages/knitr/knitr.pdf)" and "[ggiraph][1]" packages. 

```{r required packages, warning=FALSE, include=FALSE}
### Install/load required packages
#List of R packages required for this analysis:
required_packages <- c("ggplot2", "stringr", "knitr", "dplyr", "ggiraph")
#Install required_packages:
new.packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#Load required_packages:
lapply(required_packages, require, character.only = TRUE)

#Set decimal points and disable scientific notation
options(digits=3, scipen=999) 
```

###Use the same fake revenue data as my last blog post, but this time add a column for State Total Revenue ($) per year:
```{r setup dataset, warning=FALSE}

#Make up some fake data
df<-data_frame(state=rep(c("New South Wales", 
                 "Victoria", 
                 "Queensland",
                 "Western Australia",
                 "South Australia",
                 "Tasmania"), 36)) %>%
    group_by(state) %>%
    mutate(year=c(rep(2012, 9), rep(2013,9),rep(2014, 9),rep(2015, 9))) %>%
    group_by(state, year) %>%
    mutate(`store ID` = str_c("shop_#",as.character(seq_along(state)))) %>%
    group_by(state, year, `store ID`) %>%
    mutate(`Revenue ($)` =  ifelse(state=="New South Wales", sample(x=c(100000:900000), 1),
                            ifelse(state=="Victoria", sample(x=c(100000:700000), 1),
                            ifelse(state=="Queensland", sample(x=c(100000:500000), 1),
                            ifelse(state=="Western Australia",sample(x=c(10000:200000), 1),
                            ifelse(state=="South Australia",sample(x=c(10000:90000), 1),       
                            ifelse(state=="Tasmania", sample(x=c(10000:200000), 1), NA))))))) %>%
    group_by(state, year) %>%
    mutate(`State Total Revenue ($)` = sum(`Revenue ($)`))
    

```

####Peak at the first 6 rows of the fake revenue data:
```{r Peak at dataset, warning=FALSE}
kable(head(df))

```


###Get the same ggplot from my last blog most, but this time, add user reactivity using ggiraph... 
####by replacing 'geom_boxplot' and 'geom_point' (from the plot in my last blog post) with 'geom_boxplot_interactive' and 'geom_point_interactive' from the ggiraph package, and then using ggiraph() to print the plot:

```{r Plot2}

#Create the text you want displayed in the tooltop for geom_point_interactive
  df$tooltip_point <- paste0(
    "<b>", df$`store ID`, "</b>",
    "<br>(", df$state, ")",
    "<br>Revenue:",
    "<br>$", df$`Revenue ($)`,
    "</span></div>"
  )  

#Create the text you want displayed in the tooltop for geom_boxplot_interactive
  df$tooltip_boxplot <- paste0(
    "<b>", df$state, "</b>",
    "<br>State Total Revenue:",
    "<br>$", df$`State Total Revenue ($)`,
    "</span></div>"
  )    


#Make plot reactive to user's mouse with 
#'geom_boxplot_interactive' and 'geom_point_interactive'
point_interactive_plot <- ggplot(df, aes(state, `Revenue ($)`)) +
        geom_boxplot_interactive(aes(colour=state, tooltip=tooltip_boxplot, data_id=tooltip_boxplot)) +
        geom_point_interactive(aes(colour=state, tooltip=tooltip_point, data_id=tooltip_point)) +
        theme(axis.title.x =  element_blank(),
              axis.text.x  =  element_blank(), 
              axis.title.y = element_text(face="bold", size=12),
              axis.text.y  = element_text(angle=0, vjust=0.5, size=11),
              legend.title = element_text(size=12, face="bold"),
              legend.text = element_text(size = 12, face = "bold"),
              plot.title = element_text(face="bold", size=14)) + 
        ggtitle("Store Revenue per State from 2012 to 2015") +
        facet_wrap(~year) 


# htmlwidget call
ggiraph(code = {print(point_interactive_plot)}, zoom_max = 5,
        tooltip_offx = 20, tooltip_offy = -10, 
        hover_css = "fill:red;",
        tooltip_opacity = 0.7,
        pointsize = 12)


```


####These plots using the ggiraph and ggplot2 packages are cool because you can hover mouse over data point and a lable appears with information about that data point or box. Notice I've made different information appear depending on whether the mouse is hovered over an individual data point or a state box-plot, using the tooltip_point and tooltip_boxplot variables I made. Also notice the "zoom_max = 5" in the ggiraph() function allows the user to mouse scroll to zoom in and zoom out on the plot.  David Gohel and Mike Bostock have done a great job with the ggiraph package! imho :)

The ggiraph package has a bunch of other interactive geom functions too:

* geom_bar_interactive
* geom_point_interactive
* geom_polygon_interactive
* geom_map_interactive
* geom_path_interactive
* geom_rect_interactive
* geom_segment_interactive
* geom_text_interactive
* geom_boxplot_interactive

[Go check it out][1]! 

[1]: https://cran.r-project.org/web/packages/ggiraph/ggiraph.pdf
[2]: http://dpnewman.com/plotly/
